On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2015-09-25 MS.tex
New version with comments from Ian, Mark and Daniel added

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 274/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate (GR) has often been seen as a central element of it's ecological strategy, and as ultimately linked to its traits. Yet, the literature is inconsistent about the empirical correlation between functional traits and GR, casting doubt on the capacity of traits to predict growth.

\item We propose that traits should influence growth in a way that depends on the plant's size. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate) using a meta-analysis of 112 studies ($>$ 500 correlations). To cpature the effects of size, we tested for a shift in the direction of correlation between GR and traits across three plant stages: juvenile, sapling and adult.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients and light for growth, and then ultimately to replace themselves with offspring. An idea that species strategy differences might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential relative growth rate (RGR) has continued to be used as a major strategy indicator within the CSR (Competitor - Stress tolerator - Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into different components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of ``trait ecology'' arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Wright:2004jb,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, a growing number of empirical studies indicates the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. Species with high SLA can deploy more canopy area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with high photosynthetic rates and high nitrogen (N) or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling relative growth rates under favourable conditions, high SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation is mostly much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The growing discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

@@

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass = SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between traits and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. juvenile, sapling and adult) as a first approximation of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

@@

\textbf{H2: The relationship between WD and growth rate is expected to be negative across all plant size}

@@

\textbf{Effect on absolute and relative growth rates}
Tree growth in diameter, height or mass can be expressed in either absolute (AGR) or relative terms (RGR). According to the growth model, the predicted relationship between growth and either SLA or SM in seedlings reverts if absolute AGR$_\textrm{mass}$ is used instead of RGR$_\textrm{mass}$ (see Appendix \ref{sec:growth}). For SLA, the correlation with RGR$_\textrm{mass}$ is positive but becomes negative with AGR$_\textrm{mass}$. For SM the correlation with RGR$_\textrm{mass}$ is negative but becomes positive with AGR$_\textrm{mass}$. For the others traits (WD, Hmax and Aarea), the predicted relationships are not affected by the growth measurement used.

@@

Each study had to meet some basic criteria to be included in our analysis. We restricted our meta-analysis to documents reporting: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size/stage of plants. Ultimately data was included from 112 different articles (starred references in bibliography). Where available, we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). To minimise publication bias via selective reporting of significant results within the published articles (e.g. a particular case of a ``failure to publish bias'' see \citealt{Jennions:2013ta}), we asked those authors reporting a non-significant correlation for additional information about the specific correlation coefficient when these were not provided.

Overall, we recorded coefficient correlations for a list of 19 traits (Fig. \ref{fig:fig1}): the five for which we had prior expectations, plus 14 others commonly used in the trait literature. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. SM and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 19 traits were captured by a search on SLA and WD.

@@

We were unable to obtain a consistent and continuous measure of plant size across all studies, thus we relied on categorical groupings by ``stage'' to test our hypotheses. The authors of the 112 articles used in the meta-analyses, reported their plants as belonging to any of fix different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. While these categories have been established according to the expertise of the authors, by compiling data across multiple studies and vegetation types, we were concerned about the risk of assigning plants of similar size to different categories. This risk turned out to be negligible: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages was relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m); there was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships between size variables plotted using data from the Biomass And Allometry Database \citep{Falster:2015}, these thresholds also align relatively well with one another (Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries. However, the group ``juvenile'' was largely overlapping with the three others, causing us to merge this category with one of the other on a case by case basis for each study.

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``juvenile'', ``sapling'', or ``adult''.  We preferred the term ``juvenile'' to ``seedling'', because for a portion of the scientific community ``seedling'' also refers to a particular experimental condition. Using the above cut-offs we were also able to assign a stage category to studies for which no stage have been distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 112 articles were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a new stage category. Others equivocal cases were caused by an ambiguity in the definition of ``adult'' used by the authors. Some plants with a large stature are not assessed as adult by the authors because they are still non reproductive \citep{King:2006dg,King:2006he}. Whereas plants with a small stature -because stunnded by frequent wind for example \citep{Stratton:2001ck}- are assessed as adult because they are reproductive. Here, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless the plant size or age. Results using the categories established by the authors are similar than the one using the standardized categories and are available in Fig. \ref{fig:figA6} model.1. In the main text, we report only results using the standardized categories.

@@

As a measure of effect size, we used the cross-species Pearson product-moment correlation coefficient $r$ between growth rate and trait, weighted by the number of species contributing to the relationship. For studies that reported another coefficient -- such as coefficient of determination, Spearman's rho rank correlation, or Kendall's tau rank correlation -- we converted these to correlation coefficients following \citet{Lajeunesse:2013tm}. As the magnitude of Pearson's correlation coefficient approaches $\pm$ 1.0, its distribution becomes skewed. Accordingly we transformed the correlation $r$ into a metric $z$ with more-desirable statistical properties. The transformation was achieved via Fisher's z-transformation,
\[ z = \frac{1}{2} \ln \frac{1+r}{1-r}, \]
which has a variance estimate of
\[ v_z= \frac{1}{n-3}, \]
where $n$ is the number of species. We weighted effect size by the number of species used to calculation the correlation, instead of the raw number of individuals, because our questions primarily concerned the correlation between growth and traits across species and not within species.

@@

We tested stage effect on traits for which $>$ 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g. experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in $z$ by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero.

@@

We tested how robust our results were to the particular decisions made in the analysis.
In partiuclar, we tested the sensitivity of our results to inclusion of studies that measured a large number of responses for the same trait. The idea is to check that our results are not generated by the dominance of one or few studies reporting several values for the same trait. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and then re-calculated the effect size. This produced a reduced dataset, where multiple responses for any given trait were condensed into a study-level average.
We also tested if the criteria used for including a study in the analysis affect our results. In addition of the chosen criteria, the experimental conditions (stress vs unstressed), and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Thus we stratified the data set into two subsets: one including only studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the ''ideal'' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage is by definition required to measure a trait such as for SM (i.e. seed stage), or Hmax (i.e. adult stage).

@@

\cite{Koricheva:2014ku} outlined 14 methodological quality criteria to ensure robust results from meta-analyses. Table \ref{tab:meta} outlines how we address each of these.  In particular, we quantified heterogeneity in effects size, we tested for publication bias and we explored temporal change in effect size (see Appendix \ref{app:supp_info_analyses}).

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4} a-e). The correlation between growth and SLA shifted from strongly positive at juvenile stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages. The relationship between growth and either SLA or SM in seedlings is expected to differ if absolute GR in mass is used instead of RGR mass. Here, for both traits growth in mass was exclusively expressed in relative terms for seedlings(results shown in the Fig. \ref{fig:figA8} and Fig \ref{fig:figA4} f-j)

For WD, Hmax and Aarea, correlations with growth were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling and juvenile stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{fig:fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

In total, we found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). The growth form, the experiment types and the growth measurement were not markedly best predictors of the data (results shown in the Fig. \ref{fig:figA6}).

The results reported above were confined to correlations across species, and under non-stressed conditions ($n = 176$). A ``complete'' dataset ($n = 216$) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This ``complete'' dataset generated similar effect sizes to those reported above from the ``ideal'' dataset, and led to similar conclusions (Fig. \ref{fig:fig3} grey points). Differences primarily concerned the confidence intervals around particular mean effect sizes in Fig. \ref{fig:fig3}. These differences in CI were not easily attributed to any single cause. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI 95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct};  these correlations unlike others used values of WD from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 8 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

Overall our study met all the quality criteria required for meta-analysis in ecology (Table \ref{tab:meta}). We showed that the impact of publication bias is probably trivial in our meta-analyses (Appendix E, Fig. \ref{fig:figA7}). The extent of the heterogeneity was high in our meta-analysis,  but substantially explained by the plant stage effect for SLA and SM, as expected (Appendix E). Finally, there was no indication of a temporal change in the effect size being reported within stages (Appendix E, Fig. \ref{fig:figA9}).

@@

Further, for five prominent traits, the patterns observed are consistent with specific predictions from a mechanistic growth model. This indicates provisionally that the mechanisms through which traits affect growth are correctly understood in the model. For SLA and for SM the meta-analysis confirms quite strongly the predicted influence of plant size on the trait-growth correlation. For WD, Aarea and Hmax results were consistent with predictions, but coverage was limited and additional studies would be helpful for some trait-growth-stage relationships.

>>>>

\begin{document}
\mstitlepage
\noindent
\parindent=1.5em
\addtolength{\parskip}{.3em}
\doublespacing
\linenumbers
\section{Summary 274/350}\label{abstract}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item A plant's growth rate (GR) has often been seen as a central element of it's ecological strategy, and as ultimately linked to its traits. Yet the literature is inconsistent about the empirical correlation between functional traits and GR, casting doubt on the capacity of traits to predict growth.

\item We propose that traits should influence growth in a way that depends on the size of individual plants. We outline mechanisms and hypotheses based on new theory, and test these predictions in tree species for five traits (specific leaf area, wood density, maximum height, seed mass and maximum assimilation rate) using a meta-analysis of 112 studies ($>$ 500 correlations). To capture the effects of size, we tested for a shift in the direction of correlation between GR and traits across three plant stages: juvenile, sapling and adult.

@@

The phrase ``plant ecological strategy'' refers to how species face common challenges of acquiring sufficient water, nutrients and light for growth, and then ultimately replacing themselves with offspring. An idea that species strategy differences might involve fast versus slow growth goes back at least to \citet{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential relative growth rate (RGR) has continued to be used as a major strategy indicator within the CSR (Competitor - Stress tolerator - Ruderal) classification (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into different components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of ``trait ecology'' arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Since then, traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

Currently, two main spectra of variation are widely cited as underpinning differences between species in vegetative growth rates. The ``leaf economic spectrum'' \citep{Wright:2004jb} and the ``wood economic spectrum'' \citep{Chave:2009iy} reflect tissue construction costs (for leaf area and for wood volume respectively), trading off with rates of tissue turnover or mortality. It has been widely expected that species with low tissue construction cost will typically have fast growth rates, at least in favourable physical environments \citep[e.g.][]{MullerLandau:2004dc,Poorter:2008iu,Chave:2009iy,Larjavaara:2010bn,Iida:2012jb,Paine:2015df}. Yet, a growing number of empirical studies indicate the correlation between traits and growth is not as consistent as that. Consider SLA as an illustration. Species with high SLA can deploy more canopy area for a given dry mass investment in leaves \citep{Poorter:1999wd, Reich:1992wm}. In addition, SLA is correlated with high photosynthetic rates and high nitrogen (N) or phosphorus concentrations in the leaf \citep{Wright:2004jb}. For seedling relative growth rates under favourable conditions, high SLA has repeatedly been found to be a strong predictor regardless of vegetation type or growth form \citep{Lambers:1992bj,Reich:1992wm,Grime:1997wm,Poorter:1999wd,Wright:1999ds}. Yet for adult plants, the correlation has mostly been much weaker or absent \citep{coomes_comparison_1998,Poorter:2008iu,Aiba:2009ft,Easdale:2009gv,Wright:2010tp}. The growing discrepancy between theoretical expectation and empirical results has led some to question whether traits are even useful for understanding growth \citep{Wright:2010tp, Paine:2015df}.

@@

In this paper, we used a meta-analytic approach on five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea, asymptotic or maximum height = Hmax, and seed mass = SM) to investigate:
\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item  Do the observed relationships conform to predictions (outlined in section ``Theoretical expectations'') derived from a recently developed mechanistic model \citep{Falster:2011ii}?
\end{enumerate}
We gathered correlation coefficients between traits and growth from studies that gave explicit information about the size or stage of the plants. While most studies on traits have not been concerned with effects of plant size, they have generally reported trait-growth correlations for identifiable stage classes. We used plant stage (i.e. juvenile, sapling and adult) as a surrogate for the effects of plant size. We focused on tree and woody species since the mechanistic model used to assess our predictions is most applicable to these growth forms.

@@

\textbf{H2: The relationship between WD and growth rate is expected to be negative across all plant sizes}

@@

\textbf{Effect on absolute and relative growth rates}
Tree growth in diameter, height or mass can be expressed in either absolute (AGR) or relative terms (RGR). According to the growth model, the predicted relationship between growth and SM in seedlings reverts if absolute AGR$_\textrm{mass}$ is used instead of RGR$_\textrm{mass}$ (see Appendix \ref{sec:growth}). The correlation with RGR$_\textrm{mass}$ is negative but becomes positive with AGR$_\textrm{mass}$. For SLA, WD, Hmax and Aarea the predicted relationships are not affected by the growth measurement used.

All the above hypotheses arise from direct mechanistic causation between traits and growth rate. It is possible that other correlations may arise indirectly. For example, seed size is somewhat correlated with height at maturity \citep{Westoby:2009ch}, and so an indirect correlation might arise between seed size and GR in large plants, even though no direct influence is expected. Nevertheless, for purposes of this paper we adopt working hypotheses arising only from direct mechanistic causation.

@@

Each study had to meet some basic criteria to be included in our analysis. We restricted our meta-analysis to documents reporting: i) the relationship between growth and trait as a correlation coefficient, ii) the sample size and the number of species used to perform the correlation, and iii) the size/stage of plants. Ultimately data was included from 112 different articles (starred references in bibliography). Where available, we also recorded the geographic coordinates, number of sites, experiment type (greenhouse, field experiment, forest, shade-house etc.), growth measure (relative or absolute growth rate measured using plant diameter, height or mass), vegetation type, life form (tree, woody or mixed with herbaceous species etc.), biological scale (interspecific, intraspecific or intragenus), and experimental treatments (e.g. fertilization, shade, salt stress). To minimise publication bias via selective reporting of significant results within the published articles (e.g. a particular case of a ``failure to publish bias'' see \citealt{Jennions:2013ta}), we asked those authors reporting a non-significant correlation for additional information about the specific correlation coefficient when this was not provided.

Overall, we recorded coefficient correlations for a list of 19 traits (Fig. \ref{fig:fig1}): the five for which we had prior expectations, plus 14 others commonly used in the trait literature. SLA, leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and Aarea are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. SM and Hmax capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. WD and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy  \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. According to the snapshot of the literature on Web of Science, 62\% of the literature for these 19 traits was captured by a search on SLA and WD.

@@

We were unable to obtain a consistent and continuous measure of plant size across all studies, thus we relied on categorical groupings by ``stage'' to test our hypotheses. The authors of the 112 articles used in this analysis reported their plants as belonging to any of five different stages: ``seedling'', ``juvenile'', ``sapling'', ``adult'', ``mixed''. While these categories have been established according to the expertise of the authors, by compiling data across multiple studies and vegetation types, we were concerned about the risk of assigning plants of similar size to different categories. This risk turned out to be negligible: overall, we found that the original thresholds in size/age used by the authors to distinguish different stages were relatively compatible (Fig. \ref{fig:fig2}). In particular, there was a clear break between seedlings and saplings in both age (at 1 yr) and height (at 0.5 m). There was also a clear break between saplings and adults in diameter (at 10 cm). According to simple allometric relationships between size variables plotted using data from the Biomass And Allometry Database \citep{Falster:2015}, the thresholds used for each stage transition align relatively well with one another across the different size measurements (age, height and diameter, Fig. \ref{fig:figA1}). Thus we adopted the above values as cut-offs for defining stage boundaries. However, the group ``juvenile'' was largely overlapping with the three others, and so we merged this category with one or the other on a case by case basis.

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``juvenile'', ``sapling'', or ``adult''.  We preferred the term ``juvenile'' to ``seedling'', because for a portion of the scientific community ``seedling'' also refers to a particular experimental condition. Using the above cut-offs we were also able to assign a stage category to studies for which no stage was distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 112 articles were ambiguous on their plant stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a particular stage category. Other equivocal cases were caused by ambiguity in the usage of ``adult'' by the authors. Some plants with a large stature were not described as adult by authors because they were still non reproductive \citep{King:2006dg,King:2006he}. Conversely plants with small stature -- because stunted by wind for example \citep{Stratton:2001ck}-- were sometimes assessed as adult when reproductive. For purposed of the present paper we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless of the plant's size or age. Results using the categories established by the authors are similar to those using the standardized categories (see Fig. \ref{fig:figA6} model.1). In the main text, we report only results using the standardized categories.

@@

To measure of effect size, we used the cross-species Pearson product-moment correlation coefficient $r$ between growth rate and trait, weighted by the number of species contributing to the relationship. For studies that reported another coefficient -- such as coefficient of determination, Spearman's rho rank correlation, or Kendall's tau rank correlation -- we converted these to correlation coefficients following \citet{Lajeunesse:2013tm}. As the magnitude of Pearson's correlation coefficient approaches $\pm$ 1.0, its distribution becomes skewed. Accordingly, we transformed the correlation $r$ into a metric $z$ with more-desirable statistical properties. The transformation was achieved via Fisher's z-transformation,
\[ z = \frac{1}{2} \ln \frac{1+r}{1-r}, \]
which has a variance estimate of
\[ v_z= \frac{1}{n-3}, \]
where $n$ is the number of species. We weighted effect size by the number of species used to calculate the correlation, instead of the raw number of individuals, because our questions primarily concerned the correlation between growth and traits across species and not within species.

@@

We tested stage effect on traits for which $>$ 15 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation \citep{Zuur:2009cfa}. This random effect allows us to account for the heterogeneity among studies (e.g. experimental type, vegetation type, error, etc.). For each trait, we established whether stage could explain variance in $z$ by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero.

@@

We tested how robust our results were to the particular decisions made in the analysis.
In particular, we tested the sensitivity of our results to inclusion of studies that measured a large number of responses for the same trait. The idea is to check that our results were not generated by the dominance of one or few studies reporting several values for the same trait. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and then re-calculated the effect size. This produced a reduced dataset, where multiple responses for any given trait were condensed into a study-level average.
We also tested whether the criteria used for including a study in the analysis influenced our results. In addition to the chosen criteria, the experimental conditions (stressed vs unstressed), and the concordance between stage used to measure trait and growth are clear criteria for assessing the quality of the data for addressing our hypotheses. Thus we stratified the data set into two subsets: one including only studies under unstressed conditions ('ideal' dataset), and the other set including studies under all conditions ('complete' dataset). We excluded from the ''ideal'' dataset studies where growth and traits were measured on plants from different stages, unless a particular stage was by definition required to measure the trait, such as for SM (i.e. seed stage), or Hmax (i.e. adult stage).

@@

\cite{Koricheva:2014ku} outlined 14 methodological quality criteria to ensure robust results from meta-analyses. Table \ref{tab:meta} outlines how we addressed each of these. In particular, we quantified heterogeneity in effect size, we tested for publication bias and we explored temporal change in effect size (see Appendix \ref{app:supp_info_analyses}).

@@

For both SLA and SM, the correlation between growth rate and the trait changed significantly with plant size in the manner expected. Model fits improved when stage was incorporated (see log likelihood ratio tests, Fig. \ref{fig:fig3} a and d; average values for the coefficients ($\bar{r}$) in Fig. \ref{fig:figA4} a-e). The correlation between growth and SLA shifted from strongly positive at juvenile stage $\bar{r}=0.6 \pm 0.08$, Mean $\pm$ SD) to not significantly different from zero at adult stage ($\bar{r}= -0.01 \pm 0.07$, CI 95\% overlapping zero in Fig. \ref{fig:fig3} a). The correlation between growth and SM shifted from negative at juvenile stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult stage ($\bar{r}= -0.09 \pm 0.02$, , CI 95\% overlapping zero in Fig. \ref{fig:fig3} d). For both traits, correlation coefficients recorded at sapling stage were intermediate between juvenile and adult stages. The relationship between growth and SM in seedlings is expected to differ if absolute GR in mass is used instead of RGR mass. Here, for SM growth in mass was exclusively expressed in relative terms for seedlings (results shown in the Fig. \ref{fig:figA8} and Fig \ref{fig:figA4} f-j)

For WD, Hmax and Aarea, correlations with growth were also consistent with expectations, although with limitations arising from very unequal numbers of observations between growth stages for some traits. For Hmax, the expectation was for a correlation with growth to be absent for small plants but present at intermediate and larger sizes, the mechanism being increased relative allocation to reproduction as Hmax is approached. Indeed Hmax was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$). At sapling and juvenile stages, the correlation between Hmax and growth was not significantly different from that observed at adult stage (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3} c), but also was consistent with no correlation (Fig \ref{fig:fig3} c, confidence intervals overlapping zero), due to much smaller numbers of replicate studies. We could not distinguish between these alternatives. Similarly for WD and for Aarea, results were consistent with negative (WD) and positive (Aarea) correlations with growth across all plant sizes (Fig. \ref{fig:fig3} b and e, model fit not improved when stage was incorporated). However, at the same time confidence intervals of r often spanned zero especially where there were small numbers of studies for particular growth stages.

@@

In total, we found 112 studies published between between 1983 and 2014 that met our requirements. Climate zones spanned from temperate regions to  tropics (map in Fig. \ref{fig:figA2}). Different experiment types (laboratory, field, plantation or greenhouse), growth measurements and growth form (tree, woody or across growth form) were included. Number of species used for each correlation ranged from 2 to 300 species with average 31. Given this variety, it was not surprising to find wide ranges of $r$ reported for all traits. For instance, $r$ varied from -0.98 to 0.99 for SLA, and from -0.74 to 0.74 for WD (Fig. \ref{fig:figA3}). Growth form, experiment type and growth measurement type were not strong predictors in general (Fig. \ref{fig:figA6}), and confounding with these factors is not a likely explanation fo the major results reported.

The results reported above were obtained using the "ideal" dataset, and were thereby confined to correlations across species, and under non-stressed conditions ($n = 176$). The ``complete'' dataset ($n = 216$) also included correlations measured under stressed conditions, and spanning different plant stages between the growth and trait measurements. This ``complete'' dataset generated similar effect sizes to those reported above from the ``ideal'' dataset, and led to similar conclusions (Fig. \ref{fig:fig3} grey points). Differences primarily concerned the confidence intervals around particular mean effect sizes in Fig. \ref{fig:fig3}. These differences in CI may have arisen from a variety of causes. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI 95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct}; these correlations unlike others used values of WD from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 8 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{fig:figA5}).

Overall our study met quality criteria expected for meta-analysis in ecology (Table \ref{tab:meta}). The impact of publication bias was probably trivial (Appendix E, Fig. \ref{fig:figA7}). The extent of heterogeneity was high, but was substantially explained by the plant stage effect for SLA and SM, as expected (Appendix E). Finally, there was no indication of a temporal change in the effect size being reported within stages (Appendix E, Fig. \ref{fig:figA9}).

@@

For five prominent traits, the patterns observed were consistent with specific predictions from a mechanistic growth model. This is coherent with the mechanisms through which traits affect growth being correctly understood in the model. For SLA and for SM the meta-analysis confirmed quite strongly the predicted influence of plant size on the trait-growth correlation. For WD, Aarea and Hmax results were consistent with predictions, but coverage was limited and additional studies would be helpful for some trait-growth-stage relationships.
